---
title: "BUSN 5000 Project"
subtitle: "Exploring Pay Differences between Women and Men" 
author: "First Name Last Name"
date: | 
    | Summer 2026 
    | (updated `r format(Sys.time(), '%d %b %y')`)
output: 
  ioslides_presentation:
    css: css/project_slidedeck.css
    widescreen: true
---

```{r setup, include=FALSE}
# Attach packages
library(here)          # For simple handling of relative file paths
library(tidyverse)     # To use the tidy packages
library(modelsummary)  # To make pretty tables
library(ggpmisc)       # To annotate plot fits
library(car)           # For `linearHypothesis`
library(ggrepel)       # For text annotations
library(kableExtra)    # For table formatting

# Set global options
knitr::opts_chunk$set(
  echo = TRUE, 
  warning = FALSE,
  message = FALSE
  )

# Turn off scientific notation
options(scipen=999)

options(modelsummary_html = list(
  "html" = list(
    "table.css" = "width: 100%; font-size: 8px;"  # Adjust width and font size as needed
  )
))
options(datasummary_html = list(
  "html" = list(
    "table.css" = "width: 100%; font-size: 8px;"  # Adjust width and font size as needed
  )
))

# Defensive check: confirm cpsmar_e.R has been run before this
# Rmd is knit. If out/cpsmar_e.csv is missing, the read_data
# chunk would fail with a generic "cannot open the connection"
# error; we stop here with a more helpful message.
if (!file.exists(here::here("out", "cpsmar_e.csv"))) {
  stop(
    "out/cpsmar_e.csv was not found. The project Rmd reads this ",
    "file, but it doesn't exist yet. Open r/cpsmar_e.R, Select ",
    "All, click Run, then knit this Rmd. See the Main Project ",
    "chapter of the Project Guide for a refresher: ",
    "https://abbicormier.github.io/BUSN5000-project/main-project.html"
  )
}
```

<!----------------------------------------------------------------------------->

## Academic honesty statement

I have been academically honest in all of my work and will not tolerate academic dishonesty of others, consistent with [UGA's Academic Honesty Policy](https://honesty.uga.edu/Academic-Honesty-Policy/).

Sign the academic honesty statement by typing your name on the **Signature** line. 

**Signature**: 

We will not accept submissions that omit a signed Academic Honesty statement. 

<!----------------------------------------------------------------------------->
<!----------------------------------------------------------------------------->

# Introduction

<!----------------------------------------------------------------------------->

## Overview

*Insert your content here replacing these words with your own text.*

<!----------------------------------------------------------------------------->
<!----------------------------------------------------------------------------->

# Data

<!----------------------------------------------------------------------------->

## March 2024 CPS 

*Insert your content here replacing these words with your own text.*

<!----------------------------------------------------------------------------->

## March 2024 CPS Extract 

```{r read_data, eval = FALSE}
cpsmar_e <- _____(here("_____", "_____")) 
```

*Insert your content here replacing these words with your own text.*

<!----------------------------------------------------------------------------->

## Analysis sample 

```{r btl1, eval = FALSE}
cpsmar_a <- cpsmar_e %>%
  _____(
    age >= _____, 
    age <= _____,
    earnings > _____
  ) %>%
  _____(
    gender = _____(female == _____, "Female", "Male"),
    wage = _____/(_____*_____), 
    lwage = log(_____),
    Black = case_when(race==_____~1, TRUE ~ 0),
    south = case_when(region==_____~1, TRUE ~ 0),
    married = case_when((marital==_____ | marital==_____ | marital==_____)~1, TRUE ~ 0),
    age_centered = age - 23
  )
```

*Insert your content here replacing these words with your own text.*

<!----------------------------------------------------------------------------->
<!----------------------------------------------------------------------------->

# Baseline earnings distributions   

<!----------------------------------------------------------------------------->

## Plotting earnings distributions

```{r btl2, eval = FALSE}
figure1 <- ggplot(_____, aes(x = _____, group = _____, fill = _____)) +
  geom_density(alpha = 0.4) +
  labs(
    title="Figure 1. Distribution of earnings by gender",
    x="Earnings",
    y="Density"
    )+
  theme_minimal()
earnings_fvm <- cpsmar_a %>%
  group_by(_____) %>%
  summarize(avg_earnings = round(_____(_____, na.rm = TRUE),0))

avg_earnings_f <- earnings_fvm %>% 
  _____(gender == "Female") %>% 
  pull(avg_earnings) # `pull` extracts the "avg_earnings" value for "Female" from earnings_fvm, a single value since the data only record two genders.

avg_earnings_m <- earnings_fvm %>% 
  _____(gender == "Male") %>% 
  pull(avg_earnings) # `pull` extracts the "avg_earnings" value for "Male" from earnings_fvm, a single value since the data only record two genders. 
```

<!----------------------------------------------------------------------------->

## Distribution of earnings by gender

```{r btl2.5, echo=FALSE, fig.align='center', out.width='90%', eval = FALSE}
_____
```

<!----------------------------------------------------------------------------->

## Baseline comparisons 

*Insert your content here replacing these words with your own text.*

<!----------------------------------------------------------------------------->
<!----------------------------------------------------------------------------->

# The career gender gap

<!----------------------------------------------------------------------------->

## Wages and hours differences 

<div class="table-123">
```{r mi1, echo=FALSE, eval = FALSE}
datasummary(
  _____ + _____ ~ _____*(_____ + _____ + _____),
  data=_____,
  title="Table 1. Wages and hours by gender"
  )
```
</div>

<!----------------------------------------------------------------------------->

## Documenting the differences

*Insert your content here replacing these words with your own text.*

<!----------------------------------------------------------------------------->

## Plotting career log wage profiles

```{r mi2, eval = FALSE}
cef_fvm_w <- cpsmar_a %>%
  _____(_____, _____) %>%
  _____(avg_lwage = mean(_____, na.rm = TRUE))

figure2 <- ggplot(_____, aes(x = _____, y = _____, color = _____, linetype = gender, linewidth = gender)) +
  geom_point() +
  geom_line() +
  scale_linetype_manual(values = c("Female" = "longdash", "Male" = "solid")) + 
  scale_linewidth_manual(values = c("Female" = 0.7, "Male" = 0.5)) + 
  guides(linewidth = "none") +
  labs(
    title="Figure 2. Career log-wage profiles for women and men",
    x="Year",
    y="Average log wage"
    )+
  theme_minimal()
```

<!----------------------------------------------------------------------------->

## Career log wage profiles 

```{r mi2.5, echo=FALSE, fig.align='center', out.width='90%', eval = FALSE}
_____
```

<!----------------------------------------------------------------------------->

## Estimating wage differences over a career

```{r mi3, eval = FALSE}
males <- cef_fvm_w %>%
  _____(gender == "Male") %>%
  _____(avg_lwage_male = avg_lwage) %>%
  _____(-gender) 
females <- cef_fvm_w %>%
  _____(gender == "Female") %>%
  _____(avg_lwage_female = avg_lwage) %>%
  _____(-gender)

diff_fvm <- inner_join(males, females, by = "age_centered") %>%
  _____(age_centered <= 30) %>%
  _____(
    diff = avg_lwage_male - avg_lwage_female,
    age_group = cut(
      age_centered, 
      breaks = c(-1, 10, 20, 30), 
      labels = c("1-10", "11-20", "21-30"))
    ) %>%
  group_by(_____) %>%
  _____(mean_diff = mean(_____)*100) 
  
table2 <- kable(
  diff_fvm,
  digits = 2,
  col.names = c("Year Range", "Avg Pct Difference"),
  align = "cc",
  caption = "Table 2. Percent wage differences, first 30 years",
  ) %>%
  kable_styling(position = "center")
  
```

<!----------------------------------------------------------------------------->

## Evolution of the gender wage gap

<div class="table-123">
```{r mi3.5, echo=FALSE, eval = FALSE}
_____
```
</div>

<!----------------------------------------------------------------------------->

## Discussing the gender wage gap evolution

*Insert your content here replacing these words with your own text.*

<!----------------------------------------------------------------------------->
<!----------------------------------------------------------------------------->

# Explaining the gender wage gap

<!----------------------------------------------------------------------------->

## Fitting the log wage profiles 

```{r reg1, eval = FALSE}
formula <- _____ ~ _____ + I(_____^2)
figure3 <- figure2 +  
  geom_smooth(
    method = _____, 
    formula = _____, 
    aes(group = _____), 
    se = FALSE
    ) +
  stat_poly_eq(
    aes(label =  after_stat(eq.label)),
    formula = _____,
    parse = TRUE
    ) +
  labs(
    title="Figure 3. Career log-wage profiles with quadratic fits",
    x="Year",
    y="Average log wage"
    )+
  theme_minimal()+
  theme(legend.position = "bottom")
```

<!----------------------------------------------------------------------------->

## Log wage profiles with quadratic fits  

```{r reg1.5, echo=FALSE, fig.align='center', out.width='90%', eval = FALSE}
_____
```

<!----------------------------------------------------------------------------->

## Gender differences in education

<div class="table-3">
```{r ed_vars, echo=FALSE, eval = FALSE}
datasummary(
  formula = _____,
  data = cpsmar_a,
  title = "Table 3. Educational attainment by gender"
)
```
</div>

<!----------------------------------------------------------------------------->

## Gender differences in demographics

<div class="table-3">
```{r demo_vars, echo=FALSE, eval = FALSE}
datasummary(
  formula = _____,
  data = cpsmar_a,
  title = "Table 4. Demographic characteristics by gender"
)
```
</div>

<!----------------------------------------------------------------------------->

## Documenting differences in characteristics

*Insert your content here replacing these words with your own text.*

<!----------------------------------------------------------------------------->

## Controlling for education and demographic characteristics 

```{r reg2a, eval = FALSE}
singles <- cpsmar_a %>%
  _____(
    married==_____,
    child_u6==_____
    )
models <- list(
  "Baseline"      = lm(_____ ~ _____ +
                       age_centered + I(age_centered^2),
                       _____),
  "Add Education" = lm(_____ ~ _____ +
                       age_centered + I(age_centered^2) + _____ + _____ + _____,
                       _____),
  "Add Person"    = lm(_____ ~ _____ +
                       age_centered + I(age_centered^2) + _____ + _____ + _____ +
                       _____ + _____ + _____ + _____,
                       _____),
  "Add Household" = lm(_____ ~ _____ +
                       age_centered + I(age_centered^2) + _____ + _____ + _____ +
                       _____ + _____ + _____ + _____ +
                       _____ + _____,
                       _____),
  "Only Singles"  = lm(_____ ~ _____ +
                       age_centered + I(age_centered^2) + _____ + _____ + _____ +
                       _____ + _____ + _____ + _____, 
                       _____)
)
```

<!----------------------------------------------------------------------------->

## Reporting the results 

```{r reg2b, eval = FALSE}
cm <- c(
  'female'            = 'Female',
  'age_centered'      = 'Age',
  'I(age_centered^2)' = 'Age$^2$',
  '(Intercept)'       = 'Constant'
)
gm <-  tibble::tribble(
  ~raw, ~clean, ~fmt,
  "nobs", "$N$", 0,
  "r.squared", "$R^2$", 2
)
rows <- tribble(~term, ~Baseline, ~Add_Education, ~Add_Person, ~Add_Household, ~Only_Singles,
  'Education controls',   ' ',         'X',           'X',          'X',           'X',
  'Demographic controls', ' ',         ' ',           'X',          'X',           'X',
  'Household controls',   ' ',         ' ',           ' ',          'X',           'X'
)
attr(rows, 'position') <- c(9, 10, 11) # Positions where you want these rows to appear

table5 <- modelsummary(
  models,
  add_rows = rows,
  coef_map = _____,
  gof_map = _____,
  vcov = c("_____","_____","_____","_____","_____"),
  title = "Table 5. OLS estimates of the gender wage gap",
  notes = "_____ standard errors in parentheses.",
  escape = FALSE
  )
```

<!----------------------------------------------------------------------------->

## Explaining the gender wage gap

<div class="table-4">
```{r reg2.5, echo=FALSE, eval = FALSE}
_____
```
</div>

<!----------------------------------------------------------------------------->

## Documenting the findings     

*Insert your content here replacing these words with your own text.*

<!----------------------------------------------------------------------------->
<!----------------------------------------------------------------------------->

# Conclusion 

<!----------------------------------------------------------------------------->

## Summary

*Insert your content here replacing these words with your own text.*

<!----------------------------------------------------------------------------->
<!----------------------------------------------------------------------------->

# Appendix 

<!----------------------------------------------------------------------------->

## Data documentation 

<div class="appendix-code">
```{r var_doc, eval = FALSE}
# Define the variables and their descriptions
variables <- data.frame(
  Variable = c(
    "age", 
    "earnings", 
    "hours", 
    "race", 
    "marital", 
    "HSGrad", 
    "SomeColl", 
    "CollDeg", 
    "region", 
    "female", 
    "hisp", 
    "fulltime"
    ),
  Definition = c(
    "years; capped at 85",
    "_____",
    "_____",
    "respondent’s race (1 = White only, 2 = Black only, 3 = AI only, 4 = Asian only, 5 = Hawaiian/Pacific Islander only (HP), 6 = White-Black, 7 = White-AI, 8 = White-Asian, 9 = White-HP, 11 = Black-Asian, 12 = Black-HP, 13 = AI-Asian, 14 = AI-HP, 15 = Asian-HP, 16 = White-Black-AI, 17 = White-Black-Asian, 18 = White-Black-HP, 19 = White-AI-Asian, 20 = White-AI-HP, 21 = White-Asian-HP, 22 = Black-AI-Asian, 23 = White-Black-AI-Asian, 24 = White-AI-Asian-HP, 25 = White-Black-AI-Asian-HP, 25 = Other 3 race comb., 26 = Other 4 or 5 race comb.)",
    "marital status (1 = Married civilian, 2 = Married AF, 3 = Married absent, 4 = Widowed, 5 = Divorced, 6 = Separated, 7 = Never married)",
    "= 1, if _____",
    "= 1, if _____",
    "= 1, if _____",
    "household region (1 = Northeast, 2 = Midwest, 3 = South, 4 = West)",
    "= 1 if _____",
    "= 1 if Hispanic, Spanish, or Latino",
    "= 1 if _____"
  )
)
```
</div>

<!----------------------------------------------------------------------------->

## List of main variables with definitions 

This is a list of the main variables used in this project with their definitions. 

<div class="appendix-table">
```{r var_doc_table, echo=FALSE, eval = FALSE}
# Render the table
knitr::kable(variables, format = "simple", 
      col.names = c("Variable", "Definition"))
```
</div>