---
title: "BUSN 5000 :: Pre-Project Exercise"
subtitle: Measuring Labor Force Participation
author: "First Name Last Name"
date: |
    | Summer 2026
    | (updated `r format(Sys.time(), '%d %b %y')`)
output:
  ioslides_presentation:
    css: css/project_slidedeck.css
    widescreen: false
urlcolor: "#004E60"  # Use hex code for Olympic blue from visual UGA guide
---

```{r setup, include=FALSE}
# Attach packages
library(readr)         # For read_csv
library(here)          # For easy file referencing in project workflows
# library(tidyverse)     # To use the tidy packages
# library(modelsummary)  # To make pretty tables
# library(ggpmisc)       # To annotate plot fits
# library(car)           # For `linearHypothesis`
# library(ggrepel)       # For text annotations
# library(kableExtra)    # For table formatting

# Set global options
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
  )

# Turn off scientific notation
options(scipen = 999)

# Defensive check: confirm LF.R has been run before this Rmd is
# knit. If out/LF.csv is missing, the next chunk would fail with
# a generic "cannot open the connection" error; we stop here with
# a more helpful message that points the student at the fix.
if (!file.exists(here::here("out", "LF.csv"))) {
  stop(
    "out/LF.csv was not found. The pre-project Rmd reads this ",
    "file, but it doesn't exist yet. Open r/LF.R, Select All, ",
    "click Run, then knit this Rmd. See the Setup chapter of ",
    "the Project Guide for a refresher: ",
    "https://abbicormier.github.io/BUSN5000-project/setup.html"
  )
}
```

## Read `LF.csv` and print its contents

```{r, eval = FALSE}
lf <- read_csv(here("out", "LF.csv"))
print(__)
```

## Document the contents of `LF.csv`

The `LF.csv` file contains the results of the labor force participation rate (LFPR) calculation using the March 2024 CPS produced by the script, `LF.R`. The LFPR (_____) is the ratio of the labor force (_____) to the population (_____). The labor force is the sum of the employed (_____) and unemployed (_____). The population is the sum of the labor force and those not in the labor force (_____). The LFPR is expressed as a percentage.
